---
title: "Job Loss in Covid 19"
author: "Julia Van Hoozer and Hannah Burt"
format: revealjs
editor: visual
page- layout: full
editor_options: 
  chunk_output_type: console
---

## Introduction

-   Investigate the correlation between job loss and Covid 19 in the United States.

-   Look at multiple factors including: gender, ethnicity and age. 

-   Compare these factors before and during Covid 19.

-   Look specifically in Tennessee at these rates compared to the US.


## Summary Statistics of Unemplpoyment Durring Covid 19
```{r}
library(tidyverse)
library(gt)

unrate_data<- read_csv("UNRATE.csv")
unrate_data <- unrate_data |> 
  mutate(observation_date = as.Date(observation_date))
```

```{r}
unrate_data <- unrate_data |> 
  mutate(
    period = case_when(
      observation_date < as.Date("2018-03-01") ~ "before",
      observation_date < as.Date("2020-10-01") ~ "early",
      observation_date < as.Date("2022-03-01") ~ "late",
      TRUE ~ "after"
    )
  )

summary_stats <- unrate_data |> 
  group_by(period) |> 
  summarize(
    minimum = min(UNRATE, na.rm = TRUE),
    Q1 = quantile(UNRATE, probs = 0.25, na.rm = TRUE),
    median = median(UNRATE, na.rm = TRUE),
    Q3 = quantile(UNRATE, probs = 0.75, na.rm = TRUE),
    maximum = max(UNRATE, na.rm = TRUE)
  )

summary_table <- summary_stats |> 
  gt() |> 
  tab_header(
    title = "Summary Statistics of Unemployment Rates",
    subtitle = "Grouped by Period"
  ) |> 
  fmt_number(
    columns = c(minimum, Q1, median, Q3, maximum),
    decimals = 2
  ) |> 
  tab_options(
    table.width = pct(100),
    table.font.size = 25 
  )

summary_table
```

## States With the Most Job Losses From COVID-19
```{r}
library(tidyverse)
library(gt)

covid_data<- read_csv("State_COVID19_Impact.csv")
covid_data <- covid_data |>
  rename(jobs_lost = 'Unemployed_2020')|>
  mutate(jobs_lost_abs = abs(jobs_lost))

```

```{r}
top_states <- covid_data|>
  arrange(desc(jobs_lost_abs))|>
  slice_head( n=10)

top_states_table <- top_states|>
  select(State, jobs_lost_abs)|>
  rename(jobs_lost = jobs_lost_abs)|>
  gt()|>
  tab_header(
    title = "States With The Highest Job Losses Durring COVID-19"
  )|>
  fmt_number(
    columns = jobs_lost, 
    decimals = 0
  )|>
  cols_label(
    State = "State", 
    jobs_lost = "Jobs Lost"
  )|>
  tab_options(
    table.width = pct(100), 
    table.font.size = 18
  )

top_states_table
```

## Unemployment Rate from 2018-2024

```{r}
library(tidyverse)
library(janitor)
library(RColorBrewer)
library(lubridate)
library(ggplot2)
library(sf)
library(geodata)
library(dplyr)
library(readr)


UNRATE<- read_csv("UNRATE.csv")
```

```{r}
UNRATE <- UNRATE %>%
  mutate(Date = as.Date(observation_date)) %>%
  arrange(Date)
```

```{r}
ggplot(UNRATE, aes(x = Date, y = UNRATE)) + 
  geom_line(linewidth = 1, color = "#2C7FB8") +
  theme_minimal() +
  labs(title = "Unemployment Rate Trend Over Time",
       x = "Date", 
       y = "Unemployment Rate (%)") +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    axis.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )
```

## Job Loss by Age, Early, Middle, and Late COVID 19

```{r}
cdc_data <- read_csv("https://data.cdc.gov/api/views/qgkx-mswu/rows.csv?accessType=DOWNLOAD")

ggplot( cdc_data |> filter( Group == "Age group" ) )+ 
  geom_col(aes(x=Subgroup, y=`Sample Size`, fill=factor(Round)), position = "dodge") + 
  labs( title = "Loss of Work Due to COVID by Age Group Throughout COVID")
```

## Unemployment Accross The US Durring COVID 19

```{r}
#| include: FALSE
library(tidyverse)
library(sf)
library(tmap)
library(geodata)

unemployment <- read_csv("State_COVID19_Impact.csv", show_col_types = FALSE) |>
  mutate( Unemployed_2020 = abs(Unemployed_2020))

tmap_options(basemap.server = "OpenStreetMap")
tmap_mode("view")
```

```{r}

us <- gadm("USA", level = 1, path = getwd())
us <- st_as_sf(us)
us <- us |> 
  mutate( NAME_1 = toupper( NAME_1 ) ) |> 
  left_join( unemployment, by = c("NAME_1" = "State")) |>
  filter (NAME_1 != "ALASKA", NAME_1 != "HAWAII")

```

```{r}
tm_shape(us) +
  tm_polygons(id = "NAME_1", fill = "Unemployed_2020", style="jenks")
```

## Unemployment Rate in 2015

```{r}
#| include: FALSE
library(tidyverse)
library(sf)
library(tmap)
library(geodata)
unemployment <- read_csv("2015 unemployment - Sheet1.csv", show_col_types = FALSE) |>
  mutate( Unemployment_Rate = abs(Unemployment_Rate))

tmap_options(basemap.server = "OpenStreetMap")
tmap_mode("plot")
```

```{r}
tn <- gadm(country = "USA", level = 2, path = getwd())
tn <- st_as_sf(tn) 
tn <- tn |>
   filter(NAME_1=="Tennessee")

tn <- tn |> 
  left_join(unemployment, by = c("NAME_2" = "County"))

map<-tm_shape(tn)+
    tm_polygons(
    id = "NAME_2", 
    fill = "Unemployment_Rate", 
    style = "jenks",
    palette = "Reds",               
    title = "Unemployment Rate in 2015(%)"  
  )

print(map)
```

## Unemployment Rate in 2020

```{r}
#| include: FALSE
library(tidyverse)
library(sf)
library(tmap)
library(geodata)

Unemployment_2020 <- read_csv("2020 unemplyment data - Sheet1.csv", show_col_types = FALSE) |>
  mutate( Unemployment_Rate = abs(Unemployment_Rate))

tmap_options(basemap.server = "OpenStreetMap")
tmap_mode("plot")
```

```{r}
tn2 <- gadm(country = "USA", level = 2, path = getwd())
tn2 <- st_as_sf(tn2) 
tn2 <- tn |>
   filter(NAME_1=="Tennessee")
 
 tn2 <- tn2 |> 
  left_join(Unemployment_2020, by = c("NAME_2" = "county"))
 
 map2<-tm_shape(tn2)+
  tm_polygons(
    id = "NAME_2", 
    fill = "Unemployment_Rate.y", 
    style = "jenks",
    palette = "Reds",               
    title = "Unemployment Rate in 2020(%)"  
  )

print(map2)
```


